Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2024-03-16 2_SSL_NE/article/article_main.tex
various edits

\begin{abstract}
  Since their introduction to the market in 1996, white light-emitting diodes (LEDs) have greatly improved in performance, efficiency, and manufacturing cost. These advancements have been crucial for reducing global carbon emissions in the lighting sector which accounts for nearly 20\%  of global electricity consumption. Understanding the extent and sources of the rapid progress in white LED technology can thus provide valuable insights for accelerating innovation in other demand-side clean energy technologies. Through cost and performance modeling, literature review, patent analysis, and expert interviews, we find that the efficiency of top-performing  warm white GaN-based  LED packages increased from 5.8\% in 2003 to 38.8\% in 2020. Over the same period, the manufacturing cost of low-to-mid-power LED packages decreased by 95.5\% from \$1.1 to \$0.05 (in 2020 USD). Technology spillovers from other sectors contributed to 8.5\% of efficiency improvements and nearly all consumer experience enhancements, playing an important role in widespread LED adoption in lighting.
\end{abstract}

@@

A rapid reduction of global carbon dioxide emissions is urgently required in order to mitigate the effects of climate change \cite{Forster2019}. According to the Net Zero Tracker Database, by the end of 2023, more than 150 countries have set or are considering setting a net zero emissions target by 2050 \cite{climateaction}. The European Union, for example, has set a target of net zero emissions by 2050. It aims to meet this goal with the help of the European Green Deal, alongside other EU and national policies \cite{eu2020green}. The United Kingdom similarly has adopted a national strategy to achieve net-zero by 2050 \cite{noauthor_ieairena_2023}.

@@

We find that three spillovers associated with the use of YAG/YGAG phosphors in LEDs played the key role in the development of the first commercial white LED lighting products, essentially enabling the solid-state lighting market and industry of today. As noted above, subsequent spillovers then had a significant effect on physical device performance, cumulatively contributing to 8.5\% of the total LED lamp efficiency improvements between 2003 and 2020, as well as nearly all key improvements in consumer experience metrics.

@@

The cost model we developed is based on a cumulative approach to yielded cost\cite{becker2001use}. In this approach, the yielded cost $C_{Y_i}$ of process step $1$ is defined as the ratio between the total cost of step 1 $C_1$ and the yield of step 1 $Y_1$. Thus, for each consecutive step, starting from $i=1$, we get
%
\begin{equation}
    C_{Y_1} = \frac{C_1}{Y_1}, \ C_{Y_2} = C_{Y_{2 \rightarrow 3}} - C_{Y_1} = \frac{C_1(1-Y_2)+C_2}{Y_1Y_2}, \ C_{Y_3}=\dots
\end{equation}
%
Notably, yielded cost per step is dependent on the step order and blind to downstream information \cite{becker2001use}. The yielded cost metric is also cumulative by definition, thus the total cumulative yielded cost is calculated as:
%
\begin{equation}
    \sum_i C_{Y_i} = \frac{\sum_i C_i}{\prod_i Y_i}
\end{equation}
%
The overall outcome of the cost model is thus the cumulative yielded manufacturing cost per LED package for each of the three years considered. In our model, this metric includes all costs associated with producing the chip, including running costs of the factory. Costs associated with research and development, administrative overhead of the manufacturer or other investment costs are not considered. We also note that the purpose of our cost model is not to give specific estimates of the LED manufacturing cost for a factory of any size, specific geographic location or total annual manufacturing volume. It instead assumes a hypothetical factory with an assumed location in the United States and associated overhead costs related to the operation of the factory. It also assumes the use of the most up-to-date equipment for that year. Even with these simplifying assumptions, the model reasonably identifies the impact that changes in single process steps can have on the total LED manufacturing cost.

Another important limitation of our cost modelling efforts is that, even though the model captures three different chip architectures in its structure, in the present study we are able to collect, estimate and present the full set of quantitative inputs and outputs only for the classical chip architecture of low- to mid-power devices. Populating the model with data for the remaining two architectures requires access to proprietary industry information, which we have not been able to get thus far.

>>>>

\begin{abstract}
  Since their introduction to the market in 1996, white light-emitting diodes (LEDs) have greatly improved in performance, efficiency, and manufacturing cost. These advancements have been crucial for reducing global carbon emissions in the lighting sector which accounts for nearly 20\%  of global electricity consumption. Understanding the extent and sources of the rapid progress in white LED technology can thus provide valuable insights for accelerating innovation in other demand-side clean energy technologies. Through cost and performance modeling, literature review, patent analysis, and expert interviews, we find that the efficiency of top-performing  warm white GaN-based  LED packages increased from 5.8\% in 2003 to 38.8\% in 2020. Over the same period, the manufacturing cost of low-to-mid-power LED packages decreased by 95.5\% from \$1.1 to \$0.05 (in 2020 USD). Technology spillovers from other sectors accounted for at least 8.5\% of efficiency improvements and nearly all consumer experience enhancements, playing an important role in widespread LED adoption in lighting.
\end{abstract}

@@

A rapid reduction of global carbon dioxide emissions is urgently required in order to mitigate the effects of climate change \cite{Forster2019}. While the Paris Agreement of 2015 requires signatory states to develop national emissions reduction strategies, these are not legally binding \cite{bodansky2016legal}. However, some countries have introduced more ambitious legislation, including the European Union, which has set a target of net zero emissions by 2050. It aims to meet this goal with the help of the European Green Deal, alongside other national policies \cite{eu2020green}.

@@

We find that three spillovers associated with the use of YAG/YGAG phosphors in LEDs played the key role in the development of the first commercial white LED lighting products, essentially enabling the solid-state lighting market and industry of today. As noted above, subsequent spillovers then had a significant effect on physical device performance, cumulatively contributing at least 8.5\% of the total LED lamp efficiency improvements between 2003 and 2020, as well as nearly all key improvements in consumer experience metrics.

@@

The cost model we developed is based on a cumulative approach to yielded cost\cite{becker2001use}. In this approach, the yielded cost $C_{Y_i}$ of process step $1$ is defined as the ratio between the total cost of step 1 $C_1$ and the yield of step 1 $Y_1$. Thus, for each consecutive step, starting from $i=1$, we get
%
\begin{equation}
    C_{Y_1} = \frac{C_1}{Y_1}, \ C_{Y_2} = C_{Y_{2 \rightarrow 3}} - C_{Y_1} = \frac{C_1(1-Y_2)+C_2}{Y_1Y_2}, \ C_{Y_3}=\dots
\end{equation}
%
Notably, yielded cost per step is dependent on the step order and blind to downstream information \cite{becker2001use}. The yielded cost metric is also cumulative by definition, thus the total cumulative yielded cost is calculated as:
%
\begin{equation}
    \sum_i C_{Y_i} = \frac{\sum_i C_i}{\prod_i Y_i}
\end{equation}
%
The overall outcome of the cost model is thus the cumulative yielded manufacturing cost per LED package for each of the three years considered. In our model, this metric includes all costs associated with producing the chip, including operating costs of the factory. Costs associated with research and development, administrative overhead, or other investment costs are not considered. We note that the purpose of the cost model is not to provide specific estimates of the LED manufacturing cost for a factory of specific size, geographic location or manufacturing capacity. It is instead intended to capture the impact of specific improvements in the manufacturing process flow on overall cost. It therefore models a hypothetical factory operating the most up-to-date equipment for each model year. Fixed cost parameters are similar to those for a semiconductor factory operating in the United States. Even with these simplifying assumptions, the model reasonably identifies the impact that changes in single process steps can have on the total LED manufacturing cost.

Another important limitation of our cost modelling efforts is that, even though the model captures three different chip architectures in its structure, in the present study we are able to collect, estimate and present the full set of quantitative inputs and outputs only for the classical chip architecture of low-to-mid-power devices. Populating the model with data for the remaining two chip architectures would require additional access to proprietary industry information.

### 2024-03-16 2_SSL_NE/article/article_supplementary.tex
various edits

% custom section designations
% https://tex.stackexchange.com/a/712523
\titleformat{\section}[display]{\large\bfseries\filright}{\bfseries Supplementary Note \thesection}{0pt}{}
\crefname{section}{Supplementary Note}{Supplementary Notes}
% custom equation designations
\renewcommand{\figurename}{Supplementary Figure}
\crefname{figure}{Supplementary Figure}{Supplementary Figures}
\renewcommand{\tablename}{Supplementary Table}
\crefname{table}{Supplementary Table}{Supplementary Tables}

@@

\setlength{\fboxsep}{10pt}
\fbox{
    \parbox{\textwidth}{
        \textit{\textsc{Supplementary Information}} for: \\
        \newline 
        M.P. Weinold$^{1,2,3}$, S. Kolesnikov$^1$, L.D. Anadon$^{1,4}$ \\
        \newline
        $^1$ Centre for Environment, Energy and Natural Resource Governance, Department of Land Economy, University of Cambridge, Cambridge, UK \\
        $^2$ Group for Energy Systems Analysis, Department of Mechanical and Process Engineering, ETH Zurich, Zurich, Switzerland \\
        $^3$ Technology Assessment Group, Paul Scherrer Institute, Villigen, Switzerland \\
        $^4$ Belfer Center for Science and International Affairs, Harvard University, Cambridge MA, USA
    }
}

@@

The LED cost model developed for this publication is a bottom-up microeconomic manufacturing cost model with process-step resolution. Within the timeframe and scope laid out in the main text of this publication, it returns the total manufacturing cost of GaN-based phosphor-converted warm white light-emitting diode packages. The model assumes production at a \textit{"virtual"} factory which employs the highest performing equipment available in each year. Performance in this context comprises throughput, associated yield and resource consumption. It assumes North American utility prices as well as labour rates, which have remained constant over the time period of interest. Compared to other parameters, these fixed costs have little impact on the overall cost, as shown in \cref{fig:sensitivity} and \cref{fig:costmodel_calibration}. The equipment used to model the manufacturing processes has been chosen from the catalogues of current market leaders in semiconductor equipment manufacturing. This choice was made based on the specialization of the equipment, as some manufacturers offer product lines aimed specifically at light-emitting diode manufacturing. When more than one type of manufacturing equipment was available, the higher volume variant was chosen. The model does not consider costs associated with research and development or those associated with the construction of manufacturing facilities. It considers market trends through their effect on manufacturing parameters.

@@

\textbf{Availability of Equipment Performance Data}

@@

The model is set up up to capture the cost structure of a \textit{"virtual"} factory employing the most up-to-date equipment in every model year. It operates at a fixed-cost basis (labor cost, electricity cost, cleanroom cost, etc.), which was - due to limitations in data availability - taken to be based on a US-based semiconductor manufacturing plant. However, this does not correspond to any \textit{real, physical} location. The majority of production capacity is presently located in Asia. There, the share of domestically produced manufacturing equipment has increased from 10\% in 2008 to 65\% in 2015 \cite{tseng2015fab}. Recent investments into factories for high-power light-emitting diodes include Osram's Kulim plant in Malaysia \cite{osram2020kulim} and Lumileds' Penang plants in Malaysia \cite{lumileds2020penang}. In the production of low-power light-emitting diodes, the world's largest producer of LED packages by volume, MLS Co. operates plants in mainland China \cite{mls2020dev}. This means that it is difficult to validate model outputs directly with the help of industry data, which (even for American and European companies) is provided only for Asia-based manufacturing lines.

@@

Because the purpose of our model is to show the impact of changes in the design and manufacturing process of devices, rather than the geographic location and associated economic circumstances, we believe this to be a minor drawback. We therefore assumed a US-based manufacturing environment, where the associated overhead costs did not change during the time period considered.

We do compare cost model data to a selection of industry data in \cref{sec:cost_model_calibration} and provide an interpretation of the differing cost structure.

@@

@@

\textbf{}
We note some differences between the results of our model and the cost structure reported or projected by the DOE. For instance, the share of the epitaxy step is consistently larger in the DOE data. This can in part be explained by our model relying on state-of-the-art equipment at a virtual U.S.-based manufacturing location, while industry might not run low-power and mid-power chip production on these, more expensive, reactors. In addition, the manufacturing lines of the majority of manufacturers are located in Asia. We also note that the share of the sapphire substrate price in the DOE data is much larger than in our model, which can in part be explained by the overestimation of the actual price of sapphire wafers in earlier projections made for 2015-2020. Finally, the relative importance of the packaging part of the manufacturing process is very similar in our model and in the DOE results in 2012. However, in the DOE projections it significantly decreases by 2020, while in our model it retains and even increases its share of the total cost. This trend has been independently confirmed by researchers and industry reports on wafer-level packaging \cite{Lee2011WPL,Xie2013,ledsmag2017WLP}, showing better performance of our cost model compared to earlier DOE model projections.

>>>>

% custom section designations
% https://tex.stackexchange.com/a/712523
\titleformat{\section}[display]{\large\bfseries\filright}{\bfseries Supplementary Note \thesection}{-5pt}{}
\crefname{section}{Supplementary Note}{Supplementary Notes}
% custom equation designations
\renewcommand{\figurename}{Supplementary Figure}
\crefname{figure}{Supplementary Figure}{Supplementary Figures}
\renewcommand{\tablename}{Supplementary Table}
\crefname{table}{Supplementary Table}{Supplementary Tables}

@@

\setlength{\fboxsep}{10pt}
\fbox{
    \parbox{\textwidth}{
        \textit{\textsc{Supplementary Information}} for: \\
        \newline 
        M.P. Weinold$^{1,2,3}$, S. Kolesnikov$^1$, L.D. Anadon$^{1,4}$ \\
        \newline
        $^1$Centre for Environment, Energy and Natural Resource Governance \\ Department of Land Economy, University of Cambridge, Cambridge, UK \\
        $^2$Group for Energy Systems Analysis, Department of Mechanical and Process Engineering \\ ETH Zurich, Zurich, Switzerland \\
        $^3$Technology Assessment Group, Paul Scherrer Institute  Villigen, Switzerland \\
        $^4$Belfer Center for Science and International Affairs, Harvard University, Cambridge MA, USA
    }
}

@@

The LED cost model developed for this publication is a bottom-up microeconomic manufacturing cost model with process-step resolution. Within the timeframe and scope laid out in the main text of this publication, it returns the total manufacturing cost of GaN-based phosphor-converted warm white light-emitting diode packages. The model assumes a hypothetical factory which employs the highest performing equipment available in each year. Performance in this context comprises throughput, associated yield and resource consumption. It assumes North American utility prices as well as labour rates, which have remained constant over the time period of interest. Compared to other parameters, these fixed costs have little impact on the overall cost, as shown in \cref{fig:sensitivity} and \cref{fig:costmodel_calibration}. The equipment used to model the manufacturing processes has been chosen from the catalogues of current market leaders in semiconductor equipment manufacturing. This choice was made based on the specialization of the equipment, as some manufacturers offer product lines aimed specifically at light-emitting diode manufacturing. When more than one type of manufacturing equipment was available, the higher volume variant was chosen. The model does not consider costs associated with research and development or those associated with the construction of manufacturing facilities. It considers market trends through their effect on manufacturing parameters.

@@

\textbf{Availability of Equipment Performance Data for all Chip Architectures}

@@

The model is set up up to capture the cost structure of a hypothetical factory employing the most up-to-date equipment in every model year. It operates at a fixed-cost basis (labor cost, electricity cost, cleanroom cost, etc.), which was - due to limitations in data availability - taken to be based on a US-based semiconductor manufacturing plant. However, this does not correspond to any \textit{real, physical} location. The majority of production capacity is presently located in Asia. There, the share of domestically produced manufacturing equipment has increased from 10\% in 2008 to 65\% in 2015 \cite{tseng2015fab}. Recent investments into factories for high-power light-emitting diodes include Osram's Kulim plant in Malaysia \cite{osram2020kulim} and Lumileds' Penang plants in Malaysia \cite{lumileds2020penang}. In the production of low-power light-emitting diodes, the world's largest producer of LED packages by volume, MLS Co. operates plants in mainland China \cite{mls2020dev}. This means that it is difficult to validate model outputs directly with the help of industry data, which (even for American and European companies) is provided only for Asia-based manufacturing lines.

@@

Because the purpose of our model is to show the impact of changes in the design and manufacturing process of devices, rather than the geographic location and associated economic circumstances, we believe this to be a minor drawback. For fixed cost parameters, we therefore assumed a US-based manufacturing environment. We do compare cost model data to a selection of industry data in \cref{sec:cost_model_calibration} and provide an interpretation of the differing cost structure.

@@

\vspace{-20mm}

@@

\textbf{}
We note some differences between the results of our model and the cost structure reported or projected by the DOE. For instance, the share of the epitaxy step is consistently larger in the DOE data. This can in part be explained by our model assuming the use of state-of-the-art equipment, while industry might not run low-power and mid-power chip production on these more expensive reactors. In addition, the manufacturing lines of the majority of manufacturers are located in Asia. We also note that the share of the sapphire substrate price in the DOE data is much larger than in our model, which can in part be explained by the overestimation of the actual price of sapphire wafers in earlier projections made for 2015-2020. Finally, the relative importance of the packaging part of the manufacturing process is very similar in our model and in the DOE results in 2012. However, in the DOE projections it significantly decreases by 2020, while in our model it retains and even increases its share of the total cost. This trend has been independently confirmed by researchers and industry reports on wafer-level packaging \cite{Lee2011WPL,Xie2013,ledsmag2017WLP}, showing better performance of our cost model compared to earlier DOE model projections.
